---
title: '\begin{center} Predictive Analytics: Application in the Credit Risk Domain \\ Case study teaching: Cheat sheet style \\ ("group project cover sheet") \end{center}'
author: "Prof. Walter S.A. Schwaiger (IMW/TU Wien)"
date: "Oct. 26, 2022 (PredictiveAnalytics2210.Rmd)"
output:
  pdf_document:
    toc: yes
    number_sections: TRUE
    #keep_tex: TRUE
    citation_package: natbib
bibliography: literature.bib
biblio-style: humannat
include-before:
  \pagebreak

---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, warning = FALSE, message = FALSE)
```

\newpage

# Contextualization: Loading and transforming data

Origins of Weight-Of-Evidence:
\cite{Good}, \cite{Kullback}

URLs of web publications:  
https://www.rdocumentation.org/packages/scorecard/versions/0.1.8  
https://rpubs.com/ngyongkad/scorecard  
https://stats.stackexchange.com/questions/419160/credit-scorecard-using-logistic-regression-on-r  


## Methodological and linguistic overview

In the credit risk domain two predicting models will be of special importance, i.e. Generalized Linear Models (glm) and Score Card Models (scm).

LVA: "Risk Model Management" (summer term by Dr. Lederer) by using the construction, calibration and validation (CCV) framework.

## Loading libraries

```{r}
library(scorecard)
```

## Loading external data

The variables are distinguished among predictor (feature, independent) variables and response (label, dependent) variables.

One response variable: creditability  
Five predictor variables: credit.amount, duration.in.month, credit.history, purpose, property

```{r}
data("germancredit") 
data.df<-germancredit[,c("creditability",
                      "credit.amount","duration.in.month","credit.history",
                      "purpose","property")]
head(data.df[,1:3])
head(data.df[,4])
head(data.df[,5])
str(data.df)
```

Hint: Consider the different primitive data types in R, i.e. numeric (num), factor (Factor), character (chr) and integer (int).

Hint: Consider the different composed data types in R, i.e. vector, matrix, array, data frame (df) and list. Theses types will be indicated in the names of the variables, e.g. data.df is a data frame that contains the data.

## Filtering data and transforming data types

Filtering for missing values, information values and identical values by using the var_filter() function

```{r}
data_f.df <- var_filter(data.df, 
                        "creditability")
str(data_f.df)
```

Hint: Consider the change of the "creditability" data type from "factor" to "integer"

***

Calculating the information value of the predictor variables for selecting "informative" variables

```{r}
iv(data.df,y="creditability")
```


## Splitting filtered data into train and test samples

Splitting the filtered data frame into a list that contains data frames for the train and the test samples 

```{r}
data_f.list <- split_df(data_f.df, 
                       "creditability")
class(data_f.list)
lapply(data_f.list,class)
lapply(data_f.list, dim)
str(data_f.list$train)
```

Hint: By default the splitting is 70/30 % for the train/test samples.It can be altered in the split_df() function by the argument "ratio=c(0.7,0.3)". The names of the splitted samples can be altered by the argument "name_dfs=c('train','test')."  

***

Generating a list for response variable: Dummy variable for defaults

```{r}
default.list = lapply(data_f.list, function(x) x$creditability)
```

\newpage

# Transforming the predictor variables 'x': WOE(x) and GRP(x)

## Weight-Of-Evidence (WOE)-Binning

Grouping (binning) predictor variables with respect to (w.r.t.) the WOE metric

### WOE-Binning of total data: bins.df

Grouping all, i.e. total data: Needed for total cross validation purposes
```{r}
bins.df = woebin(data_f.df, 
                 "creditability") 
```

Hint: The default binning method is method="width". Other methods are "frequ" that support numerical variables as well as "tree" and "chimerge" supporting both, i.e. numerical and categorical variables which are used in the optimal binning approach. 

***

### WOE-Binning of train and test data: bins.list

Grouping only train data: Needed for separated train/test analysis
```{r}
bins.list = woebin(data_f.list$train, 
                   "creditability",
                   save_breaks_list = "breaks.list") 
```
Hint: The breaks list is saved, because it is needed to build a report for the scorecard modeling.  As the break list is saved in a separate R-script file, this file has to be loaded and the break list is saved as a variable that can be used in the scorecard modeling report.

```{r}
breaks.list=list(
 credit.amount=c("1400", "4000", "5000", "9600"), 
 duration.in.month=c("8", "16", "26", "44"), 
 credit.history=c("no credits taken/ all credits paid back duly%,%all credits at this bank paid back duly", "existing credits paid back duly till now", "delay in paying off in the past", "critical account/ other credits existing (not at this bank)"), 
 purpose=c("retraining%,%car (used)%,%radio/television", "furniture/equipment", "repairs%,%car (new)%,%business", "domestic appliances%,%education%,%others"), 
 property=c("real estate", "building society savings agreement/ life insurance", "car or other, not in attribute Savings account/bonds", "unknown / no property")
 )
```


***

Plotting the bins (including bin statistics) 

```{r fig.height=2.5, fig.align='center'}
woebin_plot(bins.list)
```

## Bin-WOE transformation of predictor variables

### Bin-WOE transformation of total data: data_woe.df

Transforming total sample: Needed for cross validation purposes
```{r}
data_woe.df <- woebin_ply(data_f.df, bins.df)
```

***

### Bin-WOE transformation of train/test data: data_woe.list

Transforming splitted sample: Needed for train/test analysis

```{r}
data_woe.list <- lapply(data_f.list, 
                       function(x) woebin_ply(x, bins.list))
lapply(data_woe.list, class)
lapply(data_woe.list, dim)
head(data_woe.list$train[,1:3])
```


## Bin-Group (GRP) transformation of predictor variables

```{r}
data_grp.list = lapply(data_f.list, 
                       function(x) woebin_ply(x, bins.list, to = 'bin'))
lapply(data_grp.list, class)
lapply(data_grp.list, dim)
head(data_grp.list$train[,1:3])
head(data_grp.list$train[,4])
```

\newpage

# Generalized linear model (glm): Regressing response w.r.t. predictors

## Logistic regression w.r.t. original predictors (data_f.list$train)

###  Logistic regression: Including original information of numeric predictors only

For comparison to lm modeling see e.g. http://www.learnbymarketing.com/tutorials/explaining-the-lm-summary-in-r/

```{r}
summary(glm(creditability ~ credit.amount + duration.in.month, 
    family = binomial(), 
    data = data_f.list$train))
```

Hint: The selection of the train sample in the logistic regression model is important for being able to implement Out-of-Sample (OoS) tests for assessing the forecasting accuracy of the glm-models and the thereof derived scorecard-models.

###  Logistic regression: Including original information of numeric, categorical and character predictors

```{r}
data_f.glm <- glm(creditability ~ ., 
                  family = binomial(), 
                  data = data_f.list$train) 
head(data_f.glm$model)
```

***

Regression results
```{r}
#summary(data_f.glm)
names(summary(data_f.glm))
head(data_f.glm$coefficients)
```

***

Analysis of regression via "analyses of variances": anova()
```{r}
anova(data_f.glm)
```

***

Analysis of regression via "variance inflation factors": vif()
```{r}
vif(data_f.glm, merge_coef = TRUE)
```


## Logistic regression w.r.t. WOE-transformed predictors (data_woe.list$train)

The WOE-based logistic regression is the prefered regression approach as it delivers the most compact regression models.

###  WOE-based logistic regression: Including numeric, categorical and character predictors

```{r}
data_woe.glm <- glm(creditability ~ ., 
                    family = binomial(), 
                    data = data_woe.list$train) 
head(data_woe.glm$model)
```

***

Summary of regression: summary()  
```{r}
summary(data_woe.glm)
names(summary(data_woe.glm))
data_woe.glm$coefficients
```

***

Analyses of variance: anova()
```{r}
anova(data_woe.glm)
```

***

Variance inflation factors: vif()
```{r}
vif(data_woe.glm, merge_coef = TRUE)
```


###  WOE-based logistic regression: Automatic selection of the best AIC-model

Automatic selection of optimal regression model via the Akaike Information Criteria (AIC)
```{r}
glm_step <- step(data_woe.glm, 
                 direction="both", 
                 trace=FALSE) 
summary(eval(glm_step$call))
```

###  WOE-based logistic regression: Including numeric predictors only

Numeric predictors: Regression with predictor variables that are of numeric data type
```{r}
summary(glm(creditability ~ credit.amount_woe + duration.in.month_woe, 
                    family = binomial(), 
                    data = data_woe.list$train)) 
```

## Logistic regression w.r.t. GRP-transformed predictors (data_grp.list$train)

###  GRP-based logistic regression: Including GRP-information of numeric predictors only

Categorical regression type: Taking bin-groups as categorical variables (dummy variable representation) 
```{r}
summary(glm(creditability ~ credit.amount_bin + duration.in.month_bin, 
    family = binomial(), 
    data = data_grp.list$train))
```

###  GRP-based logistic regression: Including numeric, categorical and character predictors

```{r}
data_grp.glm <- glm(creditability ~ ., 
                    family = binomial(), 
                    data = data_grp.list$train) 
head(data_grp.glm$model)
```

***

```{r}
#summary(data_grp.glm)
names(summary(data_grp.glm))
head(data_grp.glm$coefficients)
```

***

```{r}
anova(data_grp.glm)
```

### Transforming categorical and character predictors into factors with the one_hot() function

```{r}
data_grp_1h.list<-list()
data_grp_1h.list$train <- one_hot(data_grp.list$train, 
                                  var_encode = c("credit.history_bin",
                                                 "purpose_bin","property_bin")
                                  )
data_grp_1h.list$test <- one_hot(data_grp.list$test, 
                                 var_encode = c("credit.history_bin", 
                                                "purpose_bin", "property_bin")
                                 )
names(data_grp_1h.list)
lapply(data_grp_1h.list,dim)
```

\newpage

# Building the scorecard-model (scm) and calculating scorepoints

Scorepoints are calculated by combing scorecard-model, which combines bin and glm information, with individual data

## Building the scorecard-model

Building the scorecard via bin and glm information resulting from train sample
```{r}
scorecard.scm <- scorecard(bins.list,
                       data_woe.glm)
names(scorecard.scm)
```

***

Investigating the content of the scorecard
```{r}
scorecard.scm$basepoints
scorecard.scm$credit.amount
```


## Calculating the scorepoints by combinig scorecard-model and individual data

Calculating the scorepoints (scores) 
```{r}
score.df = scorecard_ply(data.df,
                         scorecard.scm, 
                         only_total_score = FALSE)
names(score.df)
head(score.df)
```

## Calculating scorepoints for the splitted sample (train and test)

Generating a score list
```{r}
score.list <- lapply(data_f.list, 
                     function(x) scorecard_ply(x, scorecard.scm)) 
names(score.list)
```

Hint: The only_total_score = TRUE (= default argument) has to be used for providing two compatible lists for further processing.


## Report (saved spreadsheet format): Scorecard modeling

```{r eval=FALSE}
# {r include=FALSE}
y<-"creditability"
x<-c("credit.amount","duration.in.month","credit.history","purpose","property")
# breaks.list as defined in the woebin()
report(data_f.list, 
       y, 
       x, 
       breaks.list, 
       seed = NULL, 
       save_report = "Report01")
```

Hint: Generated report file is stored in xlsx-format.  
Hint: The R-chunk is not included (i.e. {r include=FALSE}) as otherwise the report would be included in the pdf-file as well. 

\newpage

# Predicting (forecasting) probabilities and scorepoints

## Probability prediction for the sub-samples (train and test)

```{r}
predProb.list <- lapply(data_woe.list, 
                        function(x) predict(data_woe.glm, 
                                            type = 'response', 
                                            x)
                        )
names(predProb.list)
```

Hint: Due to the fact that the data_woe.glm was calibrated for the train sample two different types of prediction can be destinguished, i.e. the in-sample (IS) prediction by using the train sample in the predict()-function, and the out-of-sample (OoS) prediction by using the test sample in the predict()-function.

```{r}
head(predProb.list$train) # In-Sample prediction
head(predProb.list$test)  # Out-of-Sample prediction
```

## Scorepoint prediction for the sub-samples (train and test)

The prediction of the scorepoints is alread incorported in the built scorecard.

```{r}
head(score.list$train)
head(score.list$test)
```

\newpage

# Analyzing scoring results and testing forecasting accuracy  

## Stability of score distributions: Population stability index (PSI)

fig.width=5
```{r fig.height=2.5}
psi.list <- perf_psi(score = score.list, 
                     label = default.list,
                     return_distr_dat = TRUE)
names(psi.list)
names(psi.list$dat)
head(psi.list$dat$score[,1:9])
psi.list$psi
psi.list$pic
```

Hint: More details of per_psi() function are given @
https://www.rdocumentation.org/packages/scorecard/versions/0.1.9/topics/perf_psi  

```{r eval=FALSE}
perf_psi(score, label = NULL, title = NULL, x_limits = NULL,
  x_tick_break = 50, show_plot = TRUE, seed = 186,
  return_distr_dat = FALSE)
# e.g. # x_limits = c(250, 700),
#      # x_tick_break = 50,
```


## Statistical analysis of scoring system

```{r}
gains.df <- gains_table(score = score.list, 
                        label = default.list,
                        method = "width")
```

Investigating the gains dataframe
```{r}
head(gains.df)
tail(gains.df)
```


## Cross validation of total and sub-samples

### Cross validation w.r.t. total data

```{r}
cv.list <- perf_cv(data_woe.df, 
                   y='creditability', 
                   no_folds = 5, 
                   binomial_metric = c('ks','auc','gini','r2'))
cv.list
```

### Cross validation w.r.t. train data

```{r}
cv.list01 <- perf_cv(data_woe.list$train, 
                   y='creditability', 
                   no_folds = 5, 
                   binomial_metric = c('ks','auc','gini','r2'))
cv.list01
```

## Probability prediction accuracy (single dataset): In-Sample and Out-of-Sample testing


### IS-testing of predicted probabilities via train sample
```{r fig.height=2.5}
perf_eva(pred = predProb.list$train, 
         label = data_woe.list$train$creditability, 
         title = 'train',
         show_plot=c("roc","ks"),
         confusion_matrix = TRUE) 
```

Hint: Threshold of confusion matrix relates to the probability of default.

### OoS-testing of predicted probabilities via test sample
```{r fig.height=2.5}
perf_eva(pred = predProb.list$test, 
         label = data_woe.list$test$creditability, 
         title = 'test',
         show_plot=c("roc","ks"),
         confusion_matrix = TRUE) 
```

## Scorepoint prediction accuracy (single dataset): In-Sample and Out-of-Sample testing

### IS-testing of predicted scores via train sample
```{r fig.height=2.5}
perf_eva(pred = score.list$train, 
         label = data_f.list$train$creditability, 
         title = 'train',
         binomial_metric = c("mse","rmse","logloss","r2", "ks","auc","gini"),
         show_plot=c("roc","ks"),
         confusion_matrix = TRUE)
```

Hint: Threshold of confusion matrix relates to the scores.

### OoS-testing of predicted scores via test sample
```{r fig.height=2.5}
perf_eva(pred = score.list$test, 
         label = data_f.list$test$creditability, 
         title = 'test',
         binomial_metric = c("gini","auc","r2", "rmse"),
         show_plot=c("roc","ks"),
         confusion_matrix = TRUE)
```

## Prediction accuracy (multiple dataset): IS- and OoS-Testing in one

Predicted probabilities
```{r fig.height=2.5}
perf_eva(pred = predProb.list, 
         label = default.list,
         binomial_metric = c("gini","auc","r2", "rmse"),
         show_plot=c("roc","ks"),
         confusion_matrix = TRUE)
```

Predicted scores
```{r fig.height=2.5}
perf_eva(pred = score.list, 
         label = default.list,
         binomial_metric = c("gini","auc","r2", "rmse"),
         show_plot=c("roc","ks"),
         confusion_matrix = TRUE)
```

\newpage

# Appendix

## Appendix: Essay style with formulas in LaTeX language

**Group project assignment**: Write a scholarly essay with full sentences, correct citations and LaTeX formulas. 

**Example essay style**: From a statistical perspective the transition from the $MPS$ to the VaR framework is related to switching the perspective from considering moments (parameters) of random variables, i.e. $\mu$ and $\sigma$, to considering the quantiles and corresponding probabilities of these variables. Specifically, the VaR measure specifies the risk of a random variable ($\tilde{P}$) via the threshold quantile ($VaR$) that is exceeded into the negative direction (i.e. $P \leq VaR$) with the loss probability ($\alpha$) or respectively, is exceeded into the positive direction (i.e. $P > VaR$) with the complementary probability, i.e. the confidence level ($1-\alpha$).

\begin{align} \label{eq:VaR}
\mathrm{Pr}\{ \tilde{P} \leq VaR \} = \alpha
\end{align}

In formula (\ref{eq:VaR}) the definition of the Value-at-Risk (VaR) is given.

\newpage

```{r eval=FALSE}
# Problem: No scorecards can be built w.r.t. grp-binned data
# but logistic regression works fine with grp-binned data
# usind "Probability" prediction only for analysing grp-binned data

scorecard_grp.scm <- scorecard(bins.list,
                       data_grp.glm)
names(scorecard_grp.scm)
scorecard_grp.scm$basepoints
scorecard_grp.scm$credit.amount
scorecard.scm$basepoints
scorecard.scm$credit.amount
```

\newpage

